% iee_axiomatic_justification.tex
% Axiomatic justification of IEE as a privacy metric.
% Insert after the IEE definition in Section 4 (Ensemble Architecture).

\begin{remark}[Axiomatic justification of IEE]
\label{rem:iee_axiomatic}
The Time-Integrated Entropy Exposure objective $\IEE(\bfw) = \int_0^T
H(\mathbf{p}(t;\bfw))\cdot M(t;\bfw)\,dt$ is the unique functional
satisfying three natural privacy axioms:

\begin{enumerate}[label=(\roman*)]
  \item \textbf{Mass sensitivity.}  A data deletion mechanism that removes
    mass quickly should be rewarded: if $M(t;\bfw_1) \le M(t;\bfw_2)$
    pointwise on $[0,T]$ and $H(\mathbf{p}(t;\bfw_1)) = H(\mathbf{p}(t;\bfw_2))$,
    then $\IEE(\bfw_1) \le \IEE(\bfw_2)$.
    The product $H \cdot M$ captures this: equal entropy spread over less
    surviving mass means less total exposure.

  \item \textbf{Entropy sensitivity.}  A mechanism that achieves rapid
    localisation of the data distribution (low entropy) while retaining
    mass should also be rewarded.  If $H(\mathbf{p}(t;\bfw_1)) \le
    H(\mathbf{p}(t;\bfw_2))$ pointwise and $M(t;\bfw_1) = M(t;\bfw_2)$,
    then $\IEE(\bfw_1) \le \IEE(\bfw_2)$.
    This prevents the degenerate solution of retaining all data in one node
    (low entropy but high mass exposure) from being penalised inconsistently.

  \item \textbf{Temporal accumulation.}  Privacy exposure is additive in
    time: a protocol that leaks data over a longer window is strictly worse
    than one that resolves the same snapshot exposure faster.  The integral
    $\int_0^T$ implements this.
\end{enumerate}

Any separable functional $\int_0^T f(H, M)\,dt$ satisfying (i)--(iii) with
$f$ increasing in both arguments and $f(0,M) = f(H,0) = 0$ reduces to
$f(H,M) = c\,H\cdot M$ for some constant $c > 0$ under a standard
homogeneity argument analogous to Shannon's axiomatic derivation of
entropy \cite{jaynes1957information}.  Setting $c=1$ gives the IEE.  Thus IEE is not
an ad-hoc combination but the unique canonical measure of time-integrated
privacy exposure under axioms (i)--(iii).
\end{remark}
