% Section: When is M6 better than M5?
% For insertion into Section 5 of the paper

\subsection{When Does the Ensemble Outperform the Best Single Mechanism?}
\label{sec:ensemble_value}

While M5 (social cascade) achieves the lowest IEE in topology-matched
experiments on Barab\'asi--Albert and Watts--Strogatz graphs, three
operational scenarios exist where the M6 ensemble provides a clear advantage.

\paragraph{Scenario (a): Topology-unknown deployment.}
In practice, the operator knows $n$ but may not know the graph topology
in advance. We define the blind IEE as
\begin{equation}
  \mathrm{IEE}_{\mathrm{blind}}(k) = \tfrac{1}{3}\bigl[\mathrm{IEE}_{\mathrm{ER}}(k) +
  \mathrm{IEE}_{\mathrm{BA}}(k) + \mathrm{IEE}_{\mathrm{WS}}(k)\bigr].
\end{equation}
M6 trained on a \emph{mixed} topology dataset (10 seeds each from ER, BA, WS)
achieves $\mathrm{IEE}_{\mathrm{blind}}(\mathrm{M6}) < \mathrm{IEE}_{\mathrm{blind}}(k)$
for all $k \in \{1,\ldots,5\}$, confirming the ensemble as the correct choice
under topological uncertainty.

\paragraph{Scenario (b): Evolving networks.}
When the graph rewires during simulation (preferential-attachment updates
every $\Delta t_{\mathrm{rewire}} = 1.0$ time unit), M6's advantage over M5
increases relative to the static case. The ensemble's vote aggregation adapts
to changing local dynamics without requiring retraining, whereas M5's fixed
cascade threshold loses calibration as hub structure strengthens.

\paragraph{Scenario (c): $H_c$ misspecification robustness.}
Sweeping $H_c \in [0.4 H_{\max}, 0.9 H_{\max}]$, the IEE range
$\max_c \mathrm{IEE} - \min_c \mathrm{IEE}$ is minimised by M6, confirming
the ensemble as the most robust mechanism when the critical threshold is
uncertain. This is consistent with Remark~8 of the original analysis.

\paragraph{Summary.}
We recommend M6 when: (i) topology is unknown or mixed, (ii) the network
rewires on the simulation timescale, or (iii) $H_c$ is uncertain by $\pm 25\%$.
M5 remains competitive---and computationally cheaper---when the topology is
known to be scale-free (BA) or small-world (WS) and $H_c$ is well-calibrated.
