% Section 3.4: M3 Derived Time-Varying Formulation
% Replaces the Appendix B content

\subsection{M3: Stochastic First-Passage Deletion (Derived Time-Varying)}
\label{sec:m3_derived}

The M3 mechanism uses a first-passage probability $\mathcal{P}(\tau \leq t)$
as its vote, where $\tau$ is the first time entropy crosses $H_c$.
The original formulation used the \emph{static} Fiedler value $\lambda_2$
as the crossing rate parameter. We now derive the correct time-varying
formulation from the spectral entropy growth equation.

\paragraph{Derivation.}
The spectral decay of the normalised distribution satisfies
(Corollary~\ref{cor:spectral_decay})
\begin{equation}
  \|\mathbf{p}(t) - \tfrac{1}{n}\mathbf{1}\|_2^2 \approx
  C_0 e^{-2\alpha\lambda_2 t},
\end{equation}
implying the instantaneous entropy growth rate
\begin{equation}
  \frac{dH}{dt} \approx 2\alpha\lambda_2\bigl(H_{\max} - H(t)\bigr).
\end{equation}
The \emph{effective spectral gap} governing entropy spreading at time $t$ is
therefore
\begin{equation}
  \label{eq:lambda_eff_derived}
  \lambda_{\mathrm{eff}}(t) = \alpha\lambda_2
    \frac{H_{\max} - H(t)}{H_{\max} - H_c},
\end{equation}
which correctly approaches zero as $H(t) \to H_{\max}$ (spreading saturates)
and equals $\alpha\lambda_2$ when $H(t) = H_c$ (at the trigger threshold).

\paragraph{Vote function.}
Substituting~\eqref{eq:lambda_eff_derived} into the exponential first-passage
approximation gives the derived M3 vote:
\begin{equation}
  v_3^{\mathrm{tv}}(t) = 1 - \exp\!\bigl(-\lambda_{\mathrm{eff}}(t)\cdot t\bigr).
\end{equation}
Unlike the static formulation, $v_3^{\mathrm{tv}}$ responds to the
\emph{current} entropy state rather than elapsed time alone.

\paragraph{Comparison with Appendix B heuristic.}
The time-varying approximation of Appendix~B scaled $\lambda_2$ by
$(H/H_{\max})(M/M_0)$, which conflates entropy saturation with mass decay.
The derived formula~\eqref{eq:lambda_eff_derived} separates these effects
and has a clean interpretation: the effective crossing rate is proportional
to the remaining entropy headroom.
